---
title: "Recalling Words"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(car)
library(pander)
library(DT) #You may need to run: install.packages("DT") 
library(viridis)
```

```{r, eval=FALSE}
# Play the chunk above and this one to get the data into your Console
View(Friendly)
?Friendly
```


<br />

## Background

Many teachers and other educators are interested in understanding how to best deliver new content to students. In general, they have two choices of how to do this.

1. The Meshed Approach
    * Deliver new content while simultaneously reviewing previously understood content.

2. The Before Approach
    * Deliver new content after fully reviewing previously understood content.

A study was performed to determine whether the *Meshed* or *Before* approaches to delivering content had any positive benefits on memory recall. 

<div style="padding-left:15px;">

##### <a href="javascript:showhide('uniquename')">The Experiment <span style="font-size:8pt;">(click to view)</span></a>


<div id="uniquename" style="display:none;">

Individuals were seated at a computer and shown a list of words. Words appeared on the screen one at a time, for two seconds each, until all words had been shown (40 total). After all words were shown, they were required to perform a few two-digit mathematical additions (like 15 + 25) for 15 seconds to avoid immediate memory recall of the words. They were then asked to write down as many of the 40 words as they could remember. They were given a maximum of 5.3 minutes to recall words.

The process of showing words and recalling words was repeated four times with the same list of words each time (four chances to get it right). The presentation of the first trial was the same for all treatment conditions. However, trials 2, 3, and 4 were slightly different for each treatment condition.

<div style="padding-left:15px;">

The `SFR` group (the control group) stands for Standard Free Recall. In all four trials the same list of 40 words was presented, in a random order each time.

The `Before` group also used the same 40 words during each trial. However, any words that were correctly recalled in a previous trial were presented first, or *before* the words that were not recalled in the last trial. After all the correct words were presented in random order, the non-recalled words were presented in a random order.

The `Meshed` group also used the same 40 words during each trial. However, words that were correctly recalled in a previous trial were alternated with a missed word during the next presentation order. 

</div>

The data records the number of correctly recalled words (out of the 40 possible) from the fourth trial. Results were obtained for 30 students, 10 in each of the three treatment groups: `SFR`, `Before`, and `Meshed`. 

</div>

##### <a href="javascript:showhide('uniquename2')">The Data <span style="font-size:8pt;">(click to view)</span></a>

<div id="uniquename2" style="display:none;">

The results from the study can be found in the `Friendly` data set in R after loading `library(car)`. 

Click the "Code" button to see the data.


```{r}
datatable(Friendly, options=list(lengthMenu = c(3,10,30)))
```


</div>

</div>

<br />

## Analysis

#### Hypothesis
Using the data, we answer the question, "Does the `Meshed` approach have positive benefits on recall?" In other words, is the median numeber of words recalled using the Meshed approach higher than the median for those using standard free recall (`SFR`)?
$$
  H_0: \text{difference in medians} = 0
$$

$$ 
  H_a: \text{difference in medians} > 0
$$

 The significance level will be set at 0.05.
 $$
  \alpha = 0.05
$$


#### Graphics and Numerical Summaries

The boxplots below show the number of words recalled by individuals who used the `Meshed` approached compared to those who used the control group, standard free recall (`SFR`).


```{r, boxplots}
# Remove the data from those using the Before approach
newdata <- Friendly %>% filter(condition != "Before")
newdata$condition <- as.character(newdata$condition)
boxplot(correct ~ condition, data = newdata, main="Student Memory Test Performance", ylab="Words Recalled Out of 40", xlab="Test Condition", col = viridis(2), frame=FALSE)

#ggplot version: boxplot(correct~condition, data=newdata, main="Student Memory Test Performance", ylab="Words Recalled Out of 40", xlab="Test Condition", col = c("#66e0ff66", "#ff751a66"))


```


The difference in the two groups can be seen better in the five number summary below:
```{r, summary}
# Create a numeric vector with rent prices by gender
sfrrecall <- filter(newdata, condition == "SFR")
meshedrecall <- filter(newdata, condition == "Meshed")

# Bind vectors
recallTable <- rbind(SFR =favstats(sfrrecall$correct), Meshed=favstats(meshedrecall$correct))

# Create the table
# HOW DO YOU CENTER THE TITLE?
pander(recallTable[c("min","Q1","median","mean","Q3","max","n")], caption="Summary of Words Correctly Recalled")
```

The median number of words recalled correctly by the individuals using `SFR` is 27 while the group using the `Meshed` approach had a median of 36.5. The minimum number of words recalled correctly by the `Meshed` group is 30 (nine words higher than the control group); however, 30 is an outlier. The maximum words recalled is only one number greater for the `Meshed` group--a perfect score of 40. 


#### Wilcoxson Rank Sum Test
A Wilcoxson Rank Sum Test was used to see whether the improvement in the median of the `Meshed` group is statistically significant.

```{r, warning=FALSE}
# Wilcoxson Rank Sum Test
pander(wilcox.test(correct~condition, data=newdata, mu=0, alternative="greater", conf.level=0.95))

```

The p-value is 0.051, just barely higher than the level of significance. There is insufficient evidence to reject the null hypothesis: $p = 0.05075 > \alpha$.


</div>
</div>


## Conclusion

Although the improvement seems useful when viewing the side-by-side box plots, the difference in recalled words between the control group (`SFR`) and the `Meshed` group is not statistically significant. However, there were ties in the data so the p-value is not exact. It is recommended that a study be conducted with a larger sample size.



