---
title: "Notes_Wk5"
author: "Dawn Galloway"
date: "1/30/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2022-01-30 Monday


How do we find the spread.
The square root of MSE
Here's the summary output where's the SSE
By using teh degrees of freedom and resdiuals std error you can get back to 



```{r s}
library(ggplot2)
cars.lm <- lm(dist~ speed, data=cars)
summary(cars.lm)
confint(cars.lm)

# SSE
15.38^2
*48 

#SSTO
# SSTO = SSE + SSR
R^2 = SSR/SSTO = 1- SSE= SSTO
# Have to be able to go forward and backward with regression output most in cheet sheet
```

# Go back and look at last fridays class activity

figure out how to say these

```{r}
#Faithful
plot(waiting ~ eruptions, data=faithful, pch=21, col="darkgray")
faith.lm <- lm(waiting ~ eruptions, data=faithful)
abline(faith.lm, col="darkgray", lwd=2)
abline(v=seq(1.2, 5, 0.5), lty=2, col="gray")
abline(h=seq(50,905, 10), lty=2, col="gray")

ggplot(faithful, aes(x=eruptions, y=waiting)) +
  geom_point(fill="skybule", color="darkgray", pch=21) +
  geom_smooth(method="lm", formula=y~x, se=T, color="darkgray")

View(faithful)
?faithful
# confidence is always narrow middle less on the ends it's for the line

# This code allows us to accurately (or at least with understanding of our inaccuracy) predict the time to the next eruption.
mypreds <- predict(faith.lm, data.frame(eruptions=1.967), interval="prediction")
# for prediction is for the dots
# prediction always wider it accounts for band and dots
# Notice that your prediction interval is about (not quite, but close) to 2 residual standard errors wider on each end than the confidence interval. Discuss with a peer why you think that is the case
#At xbar the lowest variability
myc <- predict(faith.lm, data.frame(eruptions=1.967), interval="confidence")
# difference between xi and xh (xh might be in the data set, it's a new x)
# MSE measures variance of the data around the line, the average


# creates a condfidence band, the true line falls in there
# but the data doesn't fall in there. We believe the average falls in 
# Who collected this, 1.967
# if I put conf on first it'll be hidden by prediction
ggplot(faithful, aes(x=eruptions, y=waiting)) +
  geom_point(fill="skyblue", color="darkgray", pch=21) +
  geom_smooth(method="lm", formula=y~x, se=T, color="darkgray") +
  geom_segment(aes(x=1.967, xend=1.967, y=mypreds[2], yend=mypreds[3], color = "Prediction", lwd=3)) + # se is standard error
  geom_segment(aes(x=1.967, xend=1.967, y=myc[2], yend=myc[3], color = "Confidence", lwd=3)) 

   
```

Confidence is for estimating the mean. Prediction is for estimating an individual. Learning when to use which takes practice. Confidence is always for mean average.

Predict Interval says the actual temp or eruption will fall between _

```{r}

```

## Wednesday



```{r s}

```
