---
title: "Predicting the Weather"
author: "Dawn Galloway"
date: "1/7/2023"
output: 
  html_document:
    theme: cerulean
    cold-folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(mosaic)
library(DT)
```

## Rexburg's Weather for Monday, January 9th



```{r echo=FALSE}
weather_hist <- data.table(
  Date = c("2022-01-25", "2022-01-26", "2022-01-27", "2022-02-27", "2022-01-28",
           "2022-03-21", "2022-03-22", "2022-04-27", "2022-04-28", "2022-05-08",
           "2022-05-09", "2022-06-19", "2022-06-20", "2022-07-24", "2022-07-25",
           "2022-08-28", "2022-08-29", "2022-09-30", "2022-10-01", "2022-11-26",
           "2022-11-27", "2022-12-23", "2022-12-24", "2023-06-27", "2023-07-27"),
  Temperature = c(21, 16, 10, 25, 37, 
           39, 50, 59, 52, 48,
           46, 66, 64, 91, 88,
           81, 88, 63, 59, 30,
           36, 16, 25, 34, 34),
  Barometer = c(30.50, 30.46, 30.89, 30.73, 30.46,
                30.50, 30.42, 29.87, 29.78, 29.56, 
                29.79, 29.83, 30.19, 29.91, 29.97,
                29.82, 30.05, 30.03, 30.07, 30.19,
                29.93, 30.36, 30.59, 30.19, 30.33)
)

ggplot(weather_hist, aes(y=Temperature, x=Barometer)) +
  geom_point() +
  geom_smooth(method="lm", se=F, forumula=y~x)
```



## Technical Details

### Data Collection

In general, temperature and barometric pressure have an inverse relationship. The temperature and barometer measurements from 25 days across all months and across a wide range of temperatures (from 21 to 91 degrees Fahrenheit) were collected from the Rexburg, Idaho, USA page at [timeanddate.com](https://www.timeanddate.com/weather/@5605242/historic?month=12&year=2022).  


```{r}
# datatable
datatable(weather_hist, options=list(lengthMenu = c(5,10,30)), extensions="Responsive")
```


We fit a linear regression model to the weather data.

$$
  \underbrace{Y_i}_\text{Temperature} = \overbrace{\beta_0}^\text{y-int} + \overbrace{\beta_1}^\text{slope} \underbrace{X_i}_\text{Barometric Pressure} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$
The null hypothesis for the analysis is that the slope is zero; there is not a real relationship between the barometric pressure and the temperature in Rexburg. The alternative hypothesis is that slope is not zero; there is in actuality a relationship between the barometric pressure and the temperature. The level of significance for the study will be 0.05.

$$
  H_0: \beta_1 = 0
$$

$$
  H_a: \beta_1 \neq 0
$$

$$
  \alpha = 0.05
$$
```{r echo=FALSE}


mylm <- lm(Temperature ~ Barometer, data=weather_hist)
summary(mylm)
predict(mylm, data.frame(Barometer=30.35))
par(mfrow=c(1,3))
plot(mylm, which = 1:2)
plot(mylm$residuals)
```

While the QQ plot and residuals versus order plot suggest the error terms are normal and independent. The residuals versus fitted values plot does not appear to have pattern which suggests the relationship may be linear. 

$$
  \underbrace{\hat{Y}_i}_\text{Mean Temperature} = \overbrace{30.6236}^\text{est. y-int} - \overbrace{0.0095}^\text{est. slope} \underbrace{X_i}_\text{Barometric Pressure}
$$

The y-intercept for the mean temperature is 30.6236. For every point of barometric pressure, the mean Temperature decreases by -0.095 degrees Fahrenheit. The p-value for the slope is less than 0.05; therefore there is sufficient evidence to reject the null hypothesis that their is no relationship between barometric pressure and temperature.

$$p =  0.000361  <  \alpha = 0.05$$